---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---

# Problem Definition

## Specifying the Question
Finding the characteristics of customer groups

## Defining the Metric for Success
Identify which individuals are most likely to click on her ads

## Understanding the context
A Kenyan entrepreneur has created an online cryptography course and would want to advertise it on her blog. She currently targets audiences originating from various countries.

## Recording the Experimental Design
1. Load , read and explore the dataset. 
2. Clean the dataset.
3. Perform Exploratory Data Analysis.
    * Univariate Analysis
    * Bivariate Analysis

## Data Relevance
This dataset is really usefull it contains details about the users that visit her blog.

# Data Sourcing

```{r}
data <- read.csv('http://bit.ly/EcommerceCustomersDataset')
```

# Check the Data
```{r}
head(data)
```

```{r}
tail(data)
```

```{r}
colnames(data)
```

```{r}
ncol(data)
```

```{r}
nrow(data)
```

```{r}
str(data)
```


```{r}
summary(data)
```

# Perform Data Cleaning
## Duplicate 
```{r}
duplicated_rows <- data[duplicated(data),]
nrow(duplicated_rows)
```

```{r}
duplicated_rows <- data[duplicated(data),]
nrow(duplicated_rows)
data <- data[!duplicated(data), ]
duplicated_rows <- data[duplicated(data),]
nrow(duplicated_rows)
```

## Missing Values
```{r}
colSums(is.na(data))
```

```{r}
data <- na.omit(data)
colSums(is.na(data))
```

## Outliers
```{r}
boxplot(data$Administrative, main = "Administrative")
boxplot(data$Administrative_Duration, main = "Administrative Duration")
boxplot(data$Informational, main = "Informational")
boxplot(data$Informational_Duration, main = "Informational_Duration")
boxplot(data$ProductRelated, main = "ProductRelated")
boxplot(data$ProductRelated_Duration, main = "ProductRelated_Duration")
boxplot(data$BounceRates, main = "BounceRates")
boxplot(data$ExitRates, main = "ExitRates")
boxplot(data$ExitRates, main = "PageValues")
```

```{r}
data2 <- data
```

```{r}
# install.packages("CatEncoders")
library(CatEncoders)
```


```{r}
lenc <- LabelEncoder.fit(data2$Month)
data2$Month <- transform(lenc, data2$Month)
unique(data2$Month)
```

```{r}
lenc <- LabelEncoder.fit(data2$VisitorType)
data2$VisitorType <- transform(lenc, data2$VisitorType)
unique(data2$VisitorType)
```

```{r}
lenc <- LabelEncoder.fit(data2$VisitorType)
data2$VisitorType <- transform(lenc, data2$VisitorType)
unique(data2$VisitorType)
```


```{r}
lenc <- LabelEncoder.fit(data2$VisitorType)
data2$VisitorType <- transform(lenc, data2$VisitorType)
unique(data2$VisitorType)
```

```{r}
data3 <- data2
data2 = subset(data2, select = -c(Revenue) )
colnames(data2)
```

# Perform Exploratory Data Analysis

## Univariate Analysis

### Mean

Administrative 
```{r}
data.administrative.mean <- mean(data$Administrative)
data.administrative.mean
```

Administrative_Duration
```{r}
data.administrative.duration.mean <- mean(data$Administrative_Duration)
data.administrative.duration.mean
```

Informational
```{r}
data.informational.mean <- mean(data$Informational)
data.informational.mean
```

Informational_Duration  
```{r}
data.informational.duration.mean <- mean(data$Informational_Duration)
data.informational.duration.mean
```

ProductRelated
```{r}
data.product.related.mean <- mean(data$ProductRelated)
data.product.related.mean
```

ProductRelated_Duration 
```{r}
data.product.related.duration.mean <- mean(data$ProductRelated_Duration)
data.product.related.duration.mean
```

BounceRates 
```{r}
data.bounce.rates.mean <- mean(data$BounceRates)
data.bounce.rates.mean
```

ExitRates
```{r}
data.exit.rates.mean <- mean(data$ExitRates)
data.exit.rates.mean
```

PageValues
```{r}
data.administrative.mean <- mean(data$Administrative)
data.administrative.mean
```

### Median

Administrative 
```{r}
data.administrative.median <- median(data$Administrative)
data.administrative.median
```

Administrative_Duration
```{r}
data.administrative.duration.median <- median(data$Administrative_Duration)
data.administrative.duration.median
```

Informational
```{r}
data.informational.median <- median(data$Informational)
data.informational.median
```

Informational_Duration  
```{r}
data.informational.duration.median <- median(data$Informational_Duration)
data.informational.duration.median
```

ProductRelated
```{r}
data.product.related.median <- median(data$ProductRelated)
data.product.related.median
```

ProductRelated_Duration 
```{r}
data.product.related.duration.median <- median(data$ProductRelated_Duration)
data.product.related.duration.median
```

BounceRates 
```{r}
data.bounce.rates.median <- median(data$BounceRates)
data.bounce.rates.median
```

ExitRates
```{r}
data.exit.rates.median <- median(data$ExitRates)
data.exit.rates.median
```

PageValues
```{r}
data.administrative.median<- median(data$Administrative)
data.administrative.median
```

### Mode
```{r}
getmode <- function(v) {
   uniqv <- unique(v)
   uniqv[which.max(tabulate(match(v, uniqv)))]
}
```

Month
```{r}
data.month.mode <- getmode(data$Month)
```

OperatingSystems
```{r}
data.operating.system.mode <- getmode(data$OperatingSystems)
```

Browser
```{r}
data.browser.mode <- getmode(data$Browser)
```

Region
```{r}
data.region.mode <- getmode(data$Region)
```

TrafficType
```{r}
data.traffic.type.mode <- getmode(data$TrafficType)
```

VisitorType
```{r}
data.visitor.type.mode <- getmode(data$VisitorType)
```

Weekend
```{r}
data.weekend.mode<- getmode(data$Weekend)
```

Revenue
```{r}
data.revenue.mode <- getmode(data$Revenue)
```


### Measures of Dispersion

#### Minimum value

Administrative 
```{r}
data.administrative.min <- min(data$Administrative)
data.administrative.min
```

Administrative_Duration
```{r}
data.administrative.duration.min <- min(data$Administrative_Duration)
data.administrative.duration.min
```

Informational
```{r}
data.informational.min <- min(data$Informational)
data.informational.min
```

Informational_Duration  
```{r}
data.informational.duration.min <- min(data$Informational_Duration)
data.informational.duration.min
```

ProductRelated
```{r}
data.product.related.min <- min(data$ProductRelated)
data.product.related.min
```

ProductRelated_Duration 
```{r}
data.product.related.duration.min <- min(data$ProductRelated_Duration)
data.product.related.duration.min
```

BounceRates 
```{r}
data.bounce.rates.min <- min(data$BounceRates)
data.bounce.rates.min
```

ExitRates
```{r}
data.exit.rates.min <- min(data$ExitRates)
data.exit.rates.min
```

PageValues
```{r}
data.administrative.max <- max(data$Administrative)
data.administrative.max
```

     

#### Maximum value

Administrative 
```{r}
data.administrative.max <- max(data$Administrative)
data.administrative.max
```

Administrative_Duration
```{r}
data.administrative.duration.max <- max(data$Administrative_Duration)
data.administrative.duration.max
```

Informational
```{r}
data.informational.max <- max(data$Informational)
data.informational.max
```

Informational_Duration  
```{r}
data.informational.duration.max <- max(data$Informational_Duration)
data.informational.duration.max
```

ProductRelated
```{r}
data.product.related.max <- max(data$ProductRelated)
data.product.related.max
```

ProductRelated_Duration 
```{r}
data.product.related.duration.max <- max(data$ProductRelated_Duration)
data.product.related.duration.max
```

BounceRates 
```{r}
data.bounce.rates.max <- max(data$BounceRates)
data.bounce.rates.max
```

ExitRates
```{r}
data.exit.rates.max <- max(data$ExitRates)
data.exit.rates.max
```

PageValues
```{r}
data.administrative.max <- max(data$Administrative)
data.administrative.max
```

#### Range

Administrative 
```{r}
data.administrative.range <- range(data$Administrative)
data.administrative.range
```

Administrative_Duration
```{r}
data.administrative.duration.range <- range(data$Administrative_Duration)
data.administrative.duration.range
```

Informational
```{r}
data.informational.range <- range(data$Informational)
data.informational.range
```

Informational_Duration  
```{r}
data.informational.duration.range <- range(data$Informational_Duration)
data.informational.duration.range
```

ProductRelated
```{r}
data.product.related.range <- range(data$ProductRelated)
data.product.related.range
```

ProductRelated_Duration 
```{r}
data.product.related.duration.range <- range(data$ProductRelated_Duration)
data.product.related.duration.range
```

BounceRates 
```{r}
data.bounce.rates.range <- range(data$BounceRates)
data.bounce.rates.range
```

ExitRates
```{r}
data.exit.rates.range <- range(data$ExitRates)
data.exit.rates.range
```

PageValues
```{r}
data.administrative.range <- range(data$Administrative)
data.administrative.range
```

#### Quantile

Administrative 
```{r}
data.administrative.quantile <- quantile(data$Administrative)
data.administrative.quantile
```

Administrative_Duration
```{r}
data.administrative.duration.quantile <- quantile(data$Administrative_Duration)
data.administrative.duration.quantile
```

Informational
```{r}
data.informational.quantile <- quantile(data$Informational)
data.informational.quantile
```

Informational_Duration  
```{r}
data.informational.duration.quantile <- quantile(data$Informational_Duration)
data.informational.duration.quantile
```

ProductRelated
```{r}
data.product.related.quantile <- quantile(data$ProductRelated)
data.product.related.quantile
```

ProductRelated_Duration 
```{r}
data.product.related.duration.quantile <- quantile(data$ProductRelated_Duration)
data.product.related.duration.quantile
```

BounceRates 
```{r}
data.bounce.rates.quantile <- quantile(data$BounceRates)
data.bounce.rates.quantile
```

ExitRates
```{r}
data.exit.rates.quantile <- quantile(data$ExitRates)
data.exit.rates.quantile
```

PageValues
```{r}
data.administrative.quantile <- quantile(data$Administrative)
data.administrative.quantile
```

#### Variance

Administrative 
```{r}
data.administrative.variance <- var(data$Administrative)
data.administrative.variance
```

Administrative_Duration
```{r}
data.administrative.duration.variance <- var(data$Administrative_Duration)
data.administrative.duration.variance
```

Informational
```{r}
data.informational.variance <- var(data$Informational)
data.informational.variance
```

Informational_Duration  
```{r}
data.informational.duration.variance <- var(data$Informational_Duration)
data.informational.duration.variance
```

ProductRelated
```{r}
data.product.related.variance <- var(data$ProductRelated)
data.product.related.variance
```

ProductRelated_Duration 
```{r}
data.product.related.duration.variance <- var(data$ProductRelated_Duration)
data.product.related.duration.variance
```

BounceRates 
```{r}
data.bounce.rates.variance <- var(data$BounceRates)
data.bounce.rates.variance
```

ExitRates
```{r}
data.exit.rates.variance <- var(data$ExitRates)
data.exit.rates.variance
```

PageValues
```{r}
data.administrative.variance <- var(data$Administrative)
data.administrative.variance
```

#### Standard Deviation

Administrative 
```{r}
data.administrative.standard.deviation <- sd(data$Administrative)
data.administrative.standard.deviation
```

Administrative_Duration
```{r}
data.administrative.duration.standard.deviation <- sd(data$Administrative_Duration)
data.administrative.duration.standard.deviation
```

Informational
```{r}
data.informational.standard.deviation <- sd(data$Informational)
data.informational.standard.deviation
```

Informational_Duration  
```{r}
data.informational.duration.standard.deviation <- sd(data$Informational_Duration)
data.informational.duration.standard.deviation
```

ProductRelated
```{r}
data.product.related.standard.deviation <- sd(data$ProductRelated)
data.product.related.standard.deviation
```

ProductRelated_Duration 
```{r}
data.product.related.duration.standard.deviation <- sd(data$ProductRelated_Duration)
data.product.related.duration.standard.deviation
```

BounceRates 
```{r}
data.bounce.rates.standard.deviation <- sd(data$BounceRates)
data.bounce.rates.standard.deviation
```

ExitRates
```{r}
data.exit.rates.standard.deviation <- sd(data$ExitRates)
data.exit.rates.standard.deviation
```

PageValues
```{r}
data.administrative.standard.deviation <- sd(data$Administrative)
data.administrative.standard.deviation
```

```{r}
data_y <- data[data$Revenue==TRUE,]
data_n <- data[data$Revenue==FALSE,]
```

```{r}
barplot(table(data$Revenue), main = "Revenue")
barplot(table(data_y$Weekend), main = "Weekend")
barplot(table(data_n$Weekend), main = "Weekend")
barplot(table(data_y$Region), main = "Region")
barplot(table(data_n$Region), main = "Region")
barplot(table(data_y$TrafficType), main = "TrafficType")
barplot(table(data_y$VisitorType), main = "VisitorType")
barplot(table(data_n$VisitorType), main = "VisitorType")
```


## Bivariate Analysis

```{r}
#install.packages("magrittr") # package installations are only needed the first time you use it
#install.packages("dplyr")    # alternative installation of the %>%
library(magrittr) # needs to be run every time you start R and want to use %>%
library(dplyr)    # alternatively, this also loads %>%
```


```{r}

names(data)[names(data) == "Administrative"] <- "adm"
names(data)[names(data) == "Administrative_Duration"] <- "adm_du"
names(data)[names(data) == "Informational"] <- "info"
names(data)[names(data) == "Informational_Duration"] <- "info_du"
names(data)[names(data) == "ProductRelated"] <- "pr"
names(data)[names(data) == "ProductRelated_Duration"] <- "pr_du"
names(data)[names(data) == "BounceRates"] <- "br"
names(data)[names(data) == "ExitRates"] <- "er"
names(data)[names(data) == "PageValues"] <- "pv"
names(data)[names(data) == "SpecialDay"] <- "sd"
```



```{r}
#cor_data <- data[, c('Administrative', 'Administrative_Duration', 'Informational', 'Informational_Duration', 'ProductRelated', 'ProductRelated_Duration', 'BounceRates', 'ExitRates', 'PageValues', 'SpecialDay')]
cor_data <- data %>% select(1:10)
res <- cor(cor_data, cor_data)
round(res, 2)
```


```{r}
#install.packages("ggcorrplot")
library(ggcorrplot)
ggcorrplot::ggcorrplot(cor(cor_data))
```


# Implement the Solution
```{r}
head(data)
```


```{r}
normalize <- function(x){
  return ((x-min(x)) / (max(x)-min(x)))
}
```


```{r}
for(i in 1:ncol(data2)) {       # for-loop over columns
  data2[, i]<- normalize(data2[, i])
}

head(data2)
```

```{r}
# install.packages("NbClust")
# install.packages("openxlsx")
#install.packages("factoextra")
library(tidyverse)
library(dplyr)
library(openxlsx)
library(NbClust)
library(factoextra)
library(readxl)
```



```{r}
data2_matrix <- data.matrix(data2)
wssplot <- function(data2_matrix, nc, seed=123){
  wss <- (nrow(data2_matrix)-1)*sum(apply(data2_matrix, 2, var))
  for (i in 2:nc){
    set.seed(seed)
    wss[i] <- sum(kmeans(data2_matrix, centers = i)$withinss)}
  plot(1:nc, wss, type = "b", xlab = "Number of Clusters", ylab = "Within groups sum of squares")
}

wssplot(data2_matrix, nc=20)
```

```{r}
fviz_nbclust(data2_matrix, kmeans, method = "wss")+
  geom_vline(xintercept = 4, linetype = 2)+
  labs(subtitle = "withinSS")
fviz_nbclust(data2_matrix, kmeans, method = "silhouette")+
  labs(subtitle = "silhouette")


```


```{r}

result <- kmeans(data2, 4) 

# Previewing the no. of records in each cluster
result$size 


```

```{r}
for(i in 1:17) {
  for (j in 1:17) {
    if (i == j) {
      break
    } else {
      plot(data2[c(i,j)], col = result$cluster)
    }
    
  }
}

```

```{r}
d <- dist(data2, method = "euclidean")
res.hc <- hclust(d, method = "ward.D2")
plot(res.hc, cex = 0.6, hang = -1, labels = data2$Region)
```


# Challenge the Solution

Giving an in depth analysis, i found Hierarchical clustering give a better representation of a customer's Character.

# Follow up Questions

The data was large but it could better if the description of the dataset was more in depth.

